---
title: "Getting Started with shiny.webawesome"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with shiny.webawesome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

`shiny.webawesome` provides an R and Shiny interface to the Web Awesome
component library.

Most component wrappers are generated from the upstream Web Awesome metadata
file `custom-elements.json`. The package aims to stay close to upstream Web
Awesome names and component APIs while adopting normal R conventions such as
snake_case argument names. The package also bundles the Web Awesome runtime it
needs, so you do not need to install Web Awesome assets separately in your
app.

Because Web Awesome lives in the browser and Shiny spans both server and
client, the package includes more than generated wrappers alone. In practice,
there are four core areas to know:

- basic wrapper usage
- Shiny bindings
- the command API
- package options

This guide introduces each one briefly and points to the longer-form
documentation for deeper coverage.

# Basic Wrapper Usage

At the simplest level, you use generated wrappers as ordinary Shiny UI
functions. Wrapper names follow the Web Awesome component names with the
`wa_` prefix, and kebab-case attributes become snake_case arguments.

```{r basic-wrapper, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- wa_page(
  title = "Basic wrappers",
  wa_container(
    class = "wa-stack",
    wa_button(
      "save_button",
      "Save",
      appearance = "filled"
    ),
    wa_card("A simple card body")
  )
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```

Two package-level helpers appear here:

- `wa_page()` creates a minimal full-page scaffold and attaches the package
  dependency once at page level
- `wa_container()` creates a plain container element that is convenient for
  layout patterns and utility-class usage

For fuller layout examples, see the `Layout Utilities` article.

# Shiny Bindings

Some components expose generated Shiny bindings. These bindings are curated to
fit Shiny's reactive model rather than mirroring every browser event emitted by
the underlying Web Component.

For example, `wa_select()` can participate directly in ordinary Shiny input
flows:

```{r bindings, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- wa_page(
  title = "Bindings",
  wa_select(
    "favorite_letter",
    wa_option("A", value = "a"),
    wa_option("B", value = "b"),
    wa_option("C", value = "c")
  ),
  verbatimTextOutput("selected_value")
)

server <- function(input, output, session) {
  output$selected_value <- renderPrint({
    input$favorite_letter
  })
}

shinyApp(ui, server)
```

The details of binding categories and reactive semantics are covered in the
`Shiny Bindings` article.

# Command API

Generated bindings and update helpers cover the most common component-side
interactions, but sometimes a browser-side property or method still needs to
be driven explicitly from the server.

For those cases, the package provides a narrow command/helper layer.

```{r command-api, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- wa_page(
  title = "Command API",
  actionButton("open_dialog", "Open dialog"),
  wa_dialog(
    "dialog",
    "Dialog body"
  )
)

server <- function(input, output, session) {
  observeEvent(input$open_dialog, {
    wa_call_method("dialog", "show", session = session)
  })
}

shinyApp(ui, server)
```

When you need app-local browser glue that is easier to express in JavaScript,
`wa_js()` lets you inject a small inline snippet and pair it with ordinary
Shiny inputs via `Shiny.setInputValue()`.

The deeper guidance for choosing between generated bindings, update helpers,
command helpers, and `wa_js()` lives in the `Command API` article.

# Package Options

The package currently documents its warning and diagnostic controls through the
package option `shiny.webawesome.warnings`, which should be a named list.

Known keys currently include:

- `missing_tree_item_id`
- `command_layer`
- `command_layer_debug`

For example, you can enable additional command-layer diagnostics during
development with:

```{r package-options, eval = FALSE}
options(
  shiny.webawesome.warnings = list(
    command_layer_debug = TRUE
  )
)
```

These options are described in more detail in the `Package Options` article and
summarized in the package help page `?shiny.webawesome`.

# Next Steps

After this guide, the most relevant longer-form docs are:

- Layout Utilities
- Shiny Bindings
- Command API
- Package Options
- Examples
