---
title: "Package Options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

`shiny.webawesome` currently uses one documented package option for runtime
warnings and diagnostics:

- `shiny.webawesome.warnings`

This option should be a named list. It allows you to suppress selected
warnings or enable additional command-layer debug output when developing or
debugging an application.

Each known key is a boolean toggle and should be set to `TRUE` or `FALSE`.

The package does not currently expose a broad option surface beyond this
warning registry.

# Warning Registry

The `shiny.webawesome.warnings` option is merged with the package's defaults at
runtime.

Known keys currently include:

- `missing_tree_item_id`
- `command_layer`
- `command_layer_debug`

## `missing_tree_item_id`

Controls the warning path used by `wa_tree()` when selected descendant
`wa-tree-item` elements do not have stable DOM `id` attributes.

The default is `TRUE`.

This is useful because the tree binding publishes selected item ids back to
Shiny. Selected items without ids cannot participate in that Shiny value.

## `command_layer`

Controls warnings emitted by the runtime command bridge used by
`wa_set_property()` and `wa_call_method()`.

The default is `TRUE`.

These warnings help surface problems such as:

- missing browser targets
- missing property names
- missing or unsupported browser methods
- unsupported command names

## `command_layer_debug`

Controls additional debug logging for the runtime command bridge.

The default is `FALSE`.

When enabled, the browser console receives debug messages for successful
command-layer operations.

# Usage Patterns

The most common pattern is to set specific keys in a named list.

For example, to enable command-layer debug logging:

```{r options-debug, eval = FALSE}
options(
  shiny.webawesome.warnings = list(
    command_layer_debug = TRUE
  )
)
```

To suppress the tree-item id warning:

```{r options-tree-warning, eval = FALSE}
options(
  shiny.webawesome.warnings = list(
    missing_tree_item_id = FALSE
  )
)
```

To suppress command-layer warnings:

```{r options-command-warning, eval = FALSE}
options(
  shiny.webawesome.warnings = list(
    command_layer = FALSE
  )
)
```

You can also set more than one key at a time:

```{r options-multiple, eval = FALSE}
options(
  shiny.webawesome.warnings = list(
    command_layer = TRUE,
    command_layer_debug = TRUE
  )
)
```

# Notes

These options are mostly relevant for:

- application developers debugging command-layer behavior
- developers working with `wa_tree()` selection ids
- maintainers or advanced users investigating runtime issues

Ordinary package usage should not require setting options in most cases.

If you are relying heavily on `command_layer` or `command_layer_debug`, the
Command API guide is the most relevant companion document.
