---
title: "Shiny Bindings"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shiny Bindings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

`shiny.webawesome` does not try to forward every browser event emitted by every
Web Awesome component into Shiny.

Instead, the package generates a curated binding surface that fits Shiny's
reactive input model. In practice, that means the package favors bindings that
represent committed, meaningful Shiny input state rather than transient browser
telemetry.

This article describes the current binding categories, the design philosophy
behind them, and representative examples.

# Binding Categories

The current generated binding surface uses three categories:

## Durable semantic values

This is the default binding model.

The binding subscribes to a supported commit/value event and then reads the
component's current durable state. Examples include:

- `wa_select()` reading its current selected value
- `wa_dialog()` exposing its current `open` state
- `wa_tree()` exposing its current selected item ids

The important idea is that Shiny receives the semantic state the user would
care about, not the raw browser event name.

## Action-style bindings

Some components are inherently action-like in Shiny. For those, the binding
uses action semantics rather than pretending the component exposes a durable
value payload.

The canonical example is `wa_button()`, whose Shiny input behaves like an
action input that invalidates on each click.

## Action-with-payload bindings

Some rare components need both:

- action invalidation semantics
- a separate companion payload input exposing the latest committed payload

The canonical current example is `wa_dropdown()`.

For `wa_dropdown(input_id = "menu", ...)`:

- `input$menu` behaves like an action counter and invalidates on every
  selection, including repeated selection of the same item
- `input$menu_value` exposes the latest selected item value

This split is useful because repeated same-item selections should still
invalidate Shiny even when the latest payload value does not change.

# Design Philosophy

The package's binding model follows a few rules:

- bindings should expose the Shiny contract, not simply mirror raw browser
  event names
- default bindings should prefer committed semantic state
- action-like controls should use action semantics only through explicit policy
- action-with-payload contracts should remain narrow and explicit

This keeps the reactive surface smaller and easier to reason about.

It also preserves a clear boundary between:

- browser-local interaction handling
- Shiny-reactive input publication

If a browser-side interaction does not belong in the package's curated binding
surface, it is usually a better fit for browser-side code or the package's
command/helper layer.

# Examples

## Example: durable semantic value

`wa_select()` is a representative durable-value binding.

```{r binding-semantic, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- webawesomePage(
  title = "Semantic binding",
  wa_select(
    "favorite_letter",
    wa_option("A", value = "a"),
    wa_option("B", value = "b"),
    wa_option("C", value = "c")
  ),
  verbatimTextOutput("selected_value")
)

server <- function(input, output, session) {
  output$selected_value <- renderPrint({
    input$favorite_letter
  })
}

shinyApp(ui, server)
```

Here, the Shiny input reflects the current selected value, not a raw event
stream.

## Example: event-driven semantic state

`wa_dialog()` is also a semantic binding, but its Shiny value is derived from
committed dialog lifecycle state rather than a conventional text or scalar
form value.

```{r binding-semantic-dialog, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- webawesomePage(
  title = "Semantic dialog state",
  actionButton("open_dialog", "Open dialog"),
  wa_dialog(
    "dialog",
    label = "Example dialog",
    "Dialog body"
  ),
  verbatimTextOutput("dialog_state")
)

server <- function(input, output, session) {
  observeEvent(input$open_dialog, {
    wa_call_method("dialog", "show", session = session)
  })

  output$dialog_state <- renderPrint({
    input$dialog
  })
}

shinyApp(ui, server)
```

Here, the binding publishes the committed semantic state of the component:
whether the dialog is currently open.

## Example: structured semantic selection

`wa_tree()` is another semantic binding, but its Shiny value is a structured
selection derived from selected descendant tree items.

```{r binding-semantic-tree, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- webawesomePage(
  title = "Semantic tree selection",
  wa_tree(
    "navigation_tree",
    selection = "multiple",
    wa_tree_item("Section A", id = "section_a"),
    wa_tree_item("Section B", id = "section_b"),
    wa_tree_item("Section C", id = "section_c")
  ),
  verbatimTextOutput("tree_value")
)

server <- function(input, output, session) {
  output$tree_value <- renderPrint({
    input$navigation_tree
  })
}

shinyApp(ui, server)
```

The Shiny value is a vector of selected descendant `wa-tree-item` ids. For
stable selection values, selectable tree items should have DOM `id`
attributes.

## Example: action-style binding

`wa_button()` is action-like in Shiny.

```{r binding-action, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- webawesomePage(
  title = "Action binding",
  wa_button("run_action", "Run"),
  verbatimTextOutput("click_count")
)

server <- function(input, output, session) {
  output$click_count <- renderPrint({
    input$run_action
  })
}

shinyApp(ui, server)
```

The input invalidates on each click in the same way a Shiny action input does.

## Example: action-with-payload binding

`wa_dropdown()` is the current canonical action-with-payload example.

```{r binding-action-payload, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- webawesomePage(
  title = "Action with payload",
  wa_dropdown(
    "menu",
    wa_dropdown_item("Alpha", value = "alpha"),
    wa_dropdown_item("Beta", value = "beta"),
    trigger = wa_button("menu_trigger", "Open menu")
  ),
  verbatimTextOutput("action_value"),
  verbatimTextOutput("payload_value")
)

server <- function(input, output, session) {
  output$action_value <- renderPrint({
    input$menu
  })

  output$payload_value <- renderPrint({
    input$menu_value
  })
}

shinyApp(ui, server)
```

In this case:

- `input$menu` is the action-style invalidator
- `input$menu_value` is the latest committed payload state

# Current Bound Components

The package currently generates Shiny bindings for the following components.

| Component | Binding category | Notes |
|---|---|---|
| `wa_button()` | action | Shiny action-style invalidation on click |
| `wa_carousel()` | semantic | durable semantic value |
| `wa_checkbox()` | semantic | durable semantic value |
| `wa_color_picker()` | semantic | durable semantic value |
| `wa_comparison()` | semantic | durable semantic value |
| `wa_details()` | semantic | durable semantic value |
| `wa_dialog()` | semantic | event-driven semantic open state |
| `wa_drawer()` | semantic | event-driven semantic open state |
| `wa_dropdown()` | action-with-payload | action invalidator plus `input$<id>_value` payload |
| `wa_input()` | semantic | durable semantic value |
| `wa_number_input()` | semantic | durable semantic value |
| `wa_radio_group()` | semantic | durable semantic value |
| `wa_rating()` | semantic | durable semantic value |
| `wa_select()` | semantic | durable semantic value |
| `wa_slider()` | semantic | durable semantic value |
| `wa_switch()` | semantic | durable semantic value |
| `wa_tab_group()` | semantic | durable semantic value |
| `wa_textarea()` | semantic | durable semantic value |
| `wa_tree()` | semantic | selected descendant item ids |

In the website build, these entries should link to the corresponding reference
pages.

# Relationship to the Command API

Bindings and the command/helper layer solve different problems.

Use generated bindings when the package already exposes the Shiny-reactive
contract you need.

Use the command/helper layer when you need browser-side property or method
interaction that falls outside the generated binding/update surface.

For more on that boundary, see the Command API guide.
