---
title: "Command API"
output: html_document
---

# Overview

`shiny.webawesome` is generator-first, but it also provides a narrow
command/helper layer for advanced cases that are outside the generated
wrapper, binding, and update-helper surface.

This layer exists for cases where you need to:

- set a live browser-side property from the Shiny server
- call a browser-side method from the Shiny server
- write a small amount of app-local browser glue in JavaScript

These helpers are intentionally narrow. They are not meant to grow into a
second per-component API alongside the generated wrapper surface.

# Design Philosophy

The package's command layer is an escape hatch, not the default way to work
with components.

Prefer:

- generated wrappers for normal UI construction
- generated bindings for curated Shiny-reactive inputs
- generated update helpers when a component already exposes a supported update
  path

Use the command layer when you need browser-side interaction that is outside
those supported generated paths.

Use `wa_js()` when the missing piece is app-local browser glue rather than a
server-to-browser command.

# Server-to-Browser Commands

The package currently exposes two user-facing server-side command helpers:

- `wa_set_property()`
- `wa_call_method()`

Internally, these use a custom Shiny message channel named
`"shiny.webawesome.command"`.

## `wa_set_property()`

`wa_set_property()` sends a one-way command from the Shiny server that assigns
a live browser-side property on the target element.

This is useful when a component property needs to change from server logic but
the package does not already provide a generated update helper for that field.

```{r command-set-property, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- wa_page(
  title = "Set property",
  actionButton("open_dialog", "Open dialog"),
  wa_dialog(
    "dialog",
    label = "Example dialog",
    "Dialog body"
  )
)

server <- function(input, output, session) {
  observeEvent(input$open_dialog, {
    wa_set_property("dialog", "open", TRUE, session = session)
  })
}

shinyApp(ui, server)
```

## `wa_call_method()`

`wa_call_method()` sends a one-way command from the Shiny server that invokes a
browser-side method on the target element.

This is useful when the component already exposes a meaningful browser-side
method and you want to trigger it directly from server logic.

```{r command-call-method, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- wa_page(
  title = "Call method",
  actionButton("show_details", "Show details"),
  actionButton("hide_details", "Hide details"),
  wa_details(
    "details",
    summary = "More information",
    "Details body"
  )
)

server <- function(input, output, session) {
  observeEvent(input$show_details, {
    wa_call_method("details", "show", session = session)
  })

  observeEvent(input$hide_details, {
    wa_call_method("details", "hide", session = session)
  })
}

shinyApp(ui, server)
```

Both helpers target elements by DOM `id`, so the component must have a stable
browser id available.

# Browser Glue

`wa_js()` serves a different purpose from the server-side command helpers.

Use it when the missing piece is a small amount of browser-local logic that is
easier to express directly in JavaScript than through a server-to-browser
command.

Typical uses include:

- listening for browser-side events
- reading live component properties in the browser
- publishing derived values back to Shiny with `Shiny.setInputValue()`

```{r command-browser-glue, eval = FALSE}
library(shiny)
library(shiny.webawesome)

ui <- wa_page(
  title = "Browser glue",
  wa_js("
    function publishDialogState() {
      const dialog = document.getElementById('dialog');

      if (!dialog ||
          !window.Shiny ||
          typeof window.Shiny.setInputValue !== 'function') {
        return;
      }

      window.Shiny.setInputValue(
        'dialog_open_state',
        dialog.open,
        { priority: 'event' }
      );
    }

    document.addEventListener('wa-show', function(event) {
      if (event.target.id === 'dialog') {
        publishDialogState();
      }
    });

    document.addEventListener('wa-after-hide', function(event) {
      if (event.target.id === 'dialog') {
        publishDialogState();
      }
    });
  "),
  wa_dialog(
    "dialog",
    label = "Example dialog",
    "Dialog body"
  ),
  verbatimTextOutput("dialog_state")
)

server <- function(input, output, session) {
  output$dialog_state <- renderPrint({
    input$dialog_open_state
  })
}

shinyApp(ui, server)
```

`wa_js()` should stay small and app-local. For larger or shared scripts,
prefer standard Shiny asset patterns.

# Diagnostics and Limits

The command layer is intentionally conservative.

It does not validate that:

- a requested property really exists on the target component
- a requested method is semantically appropriate
- arbitrary payloads will serialize into useful browser values

The package's warning/diagnostic options include:

- `command_layer`
- `command_layer_debug`

These control warning and debug output for the runtime command bridge.

For example:

```{r command-options, eval = FALSE}
options(
  shiny.webawesome.warnings = list(
    command_layer_debug = TRUE
  )
)
```

# Choosing the Right Tool

Use:

- generated wrappers for ordinary UI
- generated bindings for curated reactive inputs
- generated update helpers when available
- `wa_set_property()` for one-off live property assignment from the server
- `wa_call_method()` for one-off browser-side method calls from the server
- `wa_js()` for small browser-local logic and custom `Shiny.setInputValue()`
  publication patterns

If you find yourself building a large handwritten command or browser-glue
layer, that is usually a sign to step back and choose a different approach:
extend the supported package surface where appropriate, or move the logic into
standard Shiny/JavaScript asset patterns rather than stretching these helpers
beyond their intended scope.
